---
title: "Predicting Changes in Small Business Lending:  Fixed-Effects Versus Machine Learning"
author: "Matt Brigida"
date: "October 1, 2018"
header-includes:
   - \setlength\parindent{24pt}
abstract: "This analysis....."
always_allow_html: yes
## bibliography: references.bib 
output:
  pdf_document:
    fig_caption: yes
    highlight: tango
    keep_tex: yes
    number_sections: yes
---

## Introduction

In this analysis we'll compare the ability of deep learning and traditional fixed effects models to predict changes in small business loans.

The following code will load the libraries which we will use in this analysis.

```{r libraries}
library(plm)
library(stargazer)
library("tfestimators")
library("keras")
```

## The Data

```{r dataload}
panel <- readRDS(url("https://github.com/Matt-Brigida/FFIEC_Call_Reports/blob/master/querying_data_and_analysis/analyses/panel_data_analysis/full_panel/1_panel_with_full_quarter_date/1_one_panel_all_models/full_panel.rds?raw=true"))
```

The code below will apply filters to the data to exclude outliers.  

```{r cleaning}
cols <- c("quarter", "totSBloans_Delt", "t1_LR_lagged_1_year", "tot_SB_loans_TA_lagged_1", "ROA_lagged_1", "NPA_TA_lagged_1", "total_assets_lagged_1_year", "TD_TA_lagged_1", "african_am_ind", "hispanic_ind", "de_novo", "TETA_lagged_1_year", "post_crisis_ind", "fin_crisis_ind")

panel <- panel[, cols]

panel <- panel[complete.cases(panel), ]

panel <- panel[panel$totSBloans_Delt < 2, ]
```

<!-- add summary statistics tables -->

## Purpose and Methods

The goal of this analysis is first to predict the percent change in the amount, and number, of small business loans given a set of lagged explanatory variables.  Second, we'll consider the ability of each method to explain the drivers of the change in small business loans.

The workhorses of traditional panel data econometrics are the fixed and random effects models. In this section we will estimate a fixed effects model

## Models and Results

### Fixed-Effects Model

We estimate the parameters of the following fixed effects model:

```{r cache=TRUE, results='asis'}
FEmodel1 <- plm(totSBloans_Delt ~  t1_LR_lagged_1_year + + I(t1_LR_lagged_1_year * african_am_ind) + I(t1_LR_lagged_1_year * hispanic_ind) + TETA_lagged_1_year + tot_SB_loans_TA_lagged_1 + ROA_lagged_1 + NPA_TA_lagged_1 + I(log(panel$total_assets_lagged_1_year)) + TD_TA_lagged_1 + post_crisis_ind + fin_crisis_ind + de_novo, data = panel, model = "within", effect = "individual")

summary(FEmodel1)

FEmodel2 <- plm(totSBloans_Delt ~  t1_LR_lagged_1_year + I(t1_LR_lagged_1_year * african_am_ind) + I(t1_LR_lagged_1_year * hispanic_ind) + TETA_lagged_1_year + tot_SB_loans_TA_lagged_1 + ROA_lagged_1 + NPA_TA_lagged_1 + I(log(panel$total_assets_lagged_1_year))  + TD_TA_lagged_1 + post_crisis_ind + fin_crisis_ind + de_novo + I(log(panel$total_assets_lagged_1_year) * ROA_lagged_1), data = panel, model = "within", effect = "individual")

summary(FEmodel2)

FEmodel3 <- plm(totSBloans_Delt ~  t1_LR_lagged_1_year + I(t1_LR_lagged_1_year * african_am_ind) + I(t1_LR_lagged_1_year * hispanic_ind) + TETA_lagged_1_year + tot_SB_loans_TA_lagged_1 + ROA_lagged_1 + NPA_TA_lagged_1 + I(log(panel$total_assets_lagged_1_year)) + TD_TA_lagged_1 + post_crisis_ind + fin_crisis_ind + de_novo + I(log(panel$total_assets_lagged_1_year) * ROA_lagged_1) + I(log(panel$total_assets_lagged_1_year) * NPA_TA_lagged_1), data = panel, model = "within", effect = "individual")

summary(FEmodel3)
```

```{r fe_output_table}
stargazer(FEmodel1, FEmodel2, FEmodel3, covariate.labels = c("T1LR", "T1LR * AA", "T1LR * His", "TE", "Small Business Loans", "ROA", "NPA", "ln(TA)", "Deposits", "Post Crisis", "Fin Crisis", "De Novo", "ln(TA) * ROA", "ln(TA) * NPA"), dep.var.labels = "% Change in Amt. SB Loans", digits = 3, no.space=TRUE, header=FALSE, type='latex', omit.stat=c("LL"), title = "All Banks: Determinants of the % Change in the Amount of Small-Business Loans", intercept.bottom = TRUE, notes = "Results are from fixed-effects models with bank fixed effects, for the years 2001 through 2017.  Data are quarterly.  The dependent variable is percent change in the amount of small-business loans.  Small-Business loans are defined as the sum of commercial, industrial, and commercial real-estate loans.  All variables are lagged one year relative to the dependent variable.")

```


The model is estimated using the plm package in the R programming language.  Estimation is done via maximum likelihood.

### Deep and Wide Model

We estimate the parameters of a deep and wide neural net model.  The model is implemented in Keras and using Tensorflow as a backend.
We first load the tensorflow and keras libraries and the data.

```{r}
## Create indicator and other variables

panel$log_TA_lagged_1 <- log(panel$total_assets_lagged_1_year)
panel$aa_t1_int <- panel$african_am_ind * panel$t1_LR_lagged_1_year
panel$h_t1_int <- panel$hispanic_ind * panel$t1_LR_lagged_1_year
panel$log_TA_ROA <- log(panel$total_assets_lagged_1_year) * panel$ROA_lagged_1
panel$log_TA_NPA <- log(panel$total_assets_lagged_1_year) * panel$NPA_TA_lagged_1


## get rid of NAs

panel <- panel[complete.cases(panel), ]

## create orthogonalized total equity variable
panel$TE_orthogonal <- lm(panel$TETA_lagged_1_year ~ panel$t1_LR_lagged_1_year)$resid


## create train and test set-----
indices <- sample(1:nrow(panel), size = 0.80 * nrow(panel))
train <- panel[indices, ]
test  <- panel[-indices, ]

## ready for neural net
```

Input function

```{r}
## return an input_fn for a given subset of data
panel_input_fn <- function(data, num_epochs = 1) {
  input_fn(data,
           features = c("t1_LR_lagged_1_year", "tot_SB_loans_TA_lagged_1", "ROA_lagged_1", "NPA_TA_lagged_1", "TD_TA_lagged_1", "log_TA_lagged_1", "aa_t1_int", "h_t1_int", "log_TA_ROA", "log_TA_NPA"),
           response = "totSBloans_Delt",
           batch_size = 32,
           num_epochs = num_epochs)
}
```

```{r}
cols_linear_feature <- feature_columns(
    column_numeric("tot_SB_loans_TA_lagged_1"),
    column_numeric("NPA_TA_lagged_1"),
    column_numeric("log_TA_lagged_1"),
    column_numeric("TD_TA_lagged_1"),
    column_numeric("ROA_lagged_1")
    )

cols_dnn_feature <- feature_columns(
    column_numeric("t1_LR_lagged_1_year"),
    column_numeric("aa_t1_int"),
    column_numeric("h_t1_int"),
    column_numeric("log_TA_ROA"),
    column_numeric("log_TA_NPA")
    )
```

```{r}
model1 <- dnn_linear_combined_regressor(linear_feature_columns = cols_linear_feature, dnn_feature_columns = cols_dnn_feature, dnn_hidden_units = c(100, 50))
```

```{r message=FALSE, error=FALSE, warning=FALSE}
model1 %>% train(panel_input_fn(train, num_epochs = 10))
model1 %>% evaluate(panel_input_fn(test))
```

```{r}
obs <- panel[(dim(panel)[1] - 100):dim(panel)[1], ]
preds1 <- model1 %>% predict(panel_input_fn(obs))

true_values1 <- obs$totSBloans_Delt
predictions1 <- unlist(preds1$predictions)
```

```{r}
cor.test(unlist(preds1$predictions), obs$totSBloans_Delt)
sq_errors1 <- (true_values1 - predictions1)^2
avg_error1 <- sum(sq_errors1) / (length(sq_errors1))
rmse <- sqrt(avg_error1)
rmse
```

## Conclusion
